---
title: "Protein Protein Interaction (PPI) resources"
author: 
  - name: "Jimena Monzón Sandoval"
    orcid: 0000-0002-4190-2188
    affiliation: 
      - name: Cardiff University
        city: Cardiff
        url: www.cardiff.ac.uk
date: 'Last update: `r date()`'
execute: 
  eval: false
format: 
  html:
    theme: spacelab
---

# Resources

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

```{r message=FALSE, warning=FALSE}

library(stringr)
library(R.utils)

library(org.Hs.eg.db)
library(ensembldb)
library(EnsDb.Hsapiens.v86)

library(ggplot2)
library(ggrepel)
library(MetBrewer)
library(cowplot)

```

## Sources

Links to each database:

+ [*APID*](http://cicblade.dep.usal.es:8080/APID/init.action) Agile Protein Interactomes DataServer. Integration of known experimentally validated protein-protein physical interactions (PPIs).

+ [*BIOGRID*](https://thebiogrid.org/) Biological General Repository for Interaction Datasets. Genetic and protein interaction data from model organisms and humans.

+ [*BIOPLEX*](https://bioplex.hms.harvard.edu/). Bioplex 3.0, comprehensive experimentally derived model of the human interactome.

+ [*CORUM*](http://mips.helmholtz-muenchen.de/corum/) the COmprehensive ResoUrce of Mammalian protein complexes.

+ [*HITPREDICT*](http://www.hitpredict.org/) Physical protein-protein interactions with reliability scores.

+ [*HURI*](http://www.interactome-atlas.org/) The Human Reference Interactome.

+ [*INTACT*](https://www.ebi.ac.uk/intact/home) IntAct Molecular Interaction Database.

+ [*MINT*](https://mint.bio.uniroma2.it/) The Molecular INTeraction Database.

+ [*REACTOME*](https://reactome.org/download-data). Open-source, open access, manually curated and peer-reviewed pathway database

+ [*STRING*](https://string-db.org/). STRING is a Core Data Resource as designated by Global Biodata Coalition and ELIXIR.

```{r eval=FALSE}

dir.create(path = "./02_Datasets/")

```

## ID Mapping

+ All datasets were mapped to Ensembl Gene ID from either Entrez, Uniprot or Ensembl Protein ID.

### Entrez

```{r entrez}

# Annotations Entrez ID to Ensembl Gene ID
x <- org.Hs.egENSEMBL
# Get the entrez gene IDs that are mapped to an Ensembl ID
mapped_genes <- mappedkeys(x)
# Convert to a list
xx <- as.list(x[mapped_genes])

# Built a Dataframe with unique Entrez IDs, possible multiple Ensembl Gene IDs delimited by "||"
entrez_to_ensembl <- matrix(NA, nrow = 1, ncol = 2)
for(i in 1:length(xx)){
  if(length(unlist(xx[i])) == 1){
    gci <- cbind(names(xx[i]), unlist(xx[i]))
  } else {
    gci <- cbind(names(xx[i]), paste0(unlist(xx[i]), collapse = "||"))
  }
  entrez_to_ensembl <- rbind(entrez_to_ensembl, gci)
  rm(gci)
}
rm(i, x, mapped_genes, xx)
entrez_to_ensembl <- entrez_to_ensembl[-1, ]
rownames(entrez_to_ensembl) <- 1:nrow(entrez_to_ensembl)
colnames(entrez_to_ensembl) <- c("Entrez_ID", "Ensembl_GeneID")
entrez_to_ensembl <- data.frame(entrez_to_ensembl)

```

### Uniprot

```{r uniprot}

# Uniprot annotations from Ensembl
edb <- EnsDb.Hsapiens.v86
txs <- genes(edb, columns = c("uniprot_id"))
uniprot_to_ensembl <- mcols(txs)
colnames(uniprot_to_ensembl) <- c("Uniprot_ID", "Ensembl_GeneID")

# Remove genes that do not match any Uniprot ID
if(length(which(is.na(uniprot_to_ensembl$Uniprot_ID) == TRUE)) > 0) {
 uniprot_to_ensembl <- uniprot_to_ensembl[-which(is.na(uniprot_to_ensembl$Uniprot_ID) == TRUE), ] 
}
rm(edb, txs)

uniprot_to_ensembl <- uniprot_to_ensembl[order(uniprot_to_ensembl$Uniprot_ID), ]

uniprot_to_ensembl_unique <- aggregate.data.frame(x = uniprot_to_ensembl,
                                                  by = list(uniprot_to_ensembl$Uniprot_ID),
                                                  FUN = function(x) paste0(x, collapse = "||"))
rm(uniprot_to_ensembl)
uniprot_to_ensembl <- uniprot_to_ensembl_unique[, c("Group.1", "Ensembl_GeneID")]
colnames(uniprot_to_ensembl) <- c("Uniprot_ID", "Ensembl_GeneID")
rm(uniprot_to_ensembl_unique)

```

### Ensembl Protein ID

```{r ensprot}

# Uniprot annotations from Ensembl
edb <- EnsDb.Hsapiens.v86
txs <- genes(edb, columns = c("protein_id"))
proteinID_to_ensembl <- mcols(txs)
colnames(proteinID_to_ensembl) <- c("protein_id", "Ensembl_GeneID")

# Remove genes that do not match any protein ID
if(length(which(is.na(proteinID_to_ensembl$protein_id) == TRUE)) > 0) {
 proteinID_to_ensembl <- proteinID_to_ensembl[-which(is.na(proteinID_to_ensembl$protein_id) == TRUE), ] 
}
rm(edb, txs)

proteinID_to_ensembl <- proteinID_to_ensembl[order(proteinID_to_ensembl$protein_id), ]

proteinID_to_ensembl_unique <- aggregate.data.frame(x = proteinID_to_ensembl,
                                                  by = list(proteinID_to_ensembl$protein_id),
                                                  FUN = function(x) paste0(x, collapse = "||"))
rm(proteinID_to_ensembl)
proteinID_to_ensembl <- proteinID_to_ensembl_unique[, c("Group.1", "Ensembl_GeneID")]
colnames(proteinID_to_ensembl) <- c("Protein_ID", "Ensembl_GeneID")
rm(proteinID_to_ensembl_unique)

```


## APID

```{r eval=FALSE}

# User friendly download through
# http://cicblade.dep.usal.es:8080/APID/init.action
# Organism Homo Sapiens
# Level 0. all reported proteins pairs
# filter out inter-species interactions select "YES"
# File name 9606_noISI_Q3.txt

dir.create(path = "./02_Datasets/APID/")

```

```{r eval=FALSE}

# Apid #
db <- read.delim("./02_Datasets/APID/9606_noISI_Q3.txt", sep = "\t", 
                 quote = NULL, header = TRUE, stringsAsFactors = FALSE)

# Mapping through uniprot
db <- data.frame(db, 
                 Ensembl_Gene_Interactor_A = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$UniprotID_A, table = uniprot_to_ensembl$Uniprot_ID)],
                 Ensembl_Gene_Interactor_B = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$UniprotID_B, table = uniprot_to_ensembl$Uniprot_ID)])

# Remove interactions where either interactor cannot be mapped to Ensembl Gene IDs
nm <- unique(c(which(is.na(db$Ensembl_Gene_Interactor_A) == TRUE),
               which(is.na(db$Ensembl_Gene_Interactor_B) == TRUE)))
if(length(nm)> 0){ db <- db[-nm, ]}
rm(nm)

# Separate into interactions with a single and multiple Ensembl Gene IDs
db_m2m <- unique(c(which(str_detect(string = db$Ensembl_Gene_Interactor_A, pattern = "\\|") == TRUE),
                   which(str_detect(string = db$Ensembl_Gene_Interactor_B, pattern = "\\|") == TRUE)))

db_o2o <- setdiff(x = 1:nrow(db), y = db_m2m)
db_m2m <- db[db_m2m, ]
db_o2o <- db[db_o2o, ]

# Expand interactions with multiple ensembl gene IDs
output <- matrix(NA, nrow = 1, ncol = 2, dimnames = list("ToRemove", c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")))
for(i in c(1:nrow(db_m2m))){
  in_a <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_A[i], pattern = "\\|\\|"))
  in_b <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_B[i], pattern = "\\|\\|"))
  ab <- expand.grid("Ensembl_Gene_Interactor_A" = in_a, "Ensembl_Gene_Interactor_B" = in_b)
  output <- rbind(output, ab)
}
rm(in_a, in_b, ab, i)
output <- output[-1, ]

# Combine interactions that map a single and multiple Ensembl Gene IDs 
db_ens <- rbind(db_o2o[, c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")], 
                output)
rm(db, db_m2m, db_o2o, output)

# Remove self interactions
mark_self <- apply(db_ens, 1, FUN = function(x) length(unique(x)))
if(length(which(mark_self == 1))> 0){db_ens <- db_ens[-which(mark_self == 1), ]}
rm(mark_self)

# Re-order interactions to remove duplicates
db_ens_order <- t(apply(db_ens, 1, FUN = function(x) x[order(x)]))
db_ens <- unique(db_ens_order) 
rm(db_ens_order)

# Convert to data frame
rownames(db_ens) <- 1:nrow(db_ens)
db_ens <- as.data.frame(db_ens)

# Save as tab delimited file 
write.table(x = db_ens, 
            file = "./02_Datasets/APID/9606_noISI_Q3_EnsemblGeneID.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

```

## BIOGRID

```{r eval=FALSE}

# Download data from Biogrid by species
dir.create(path = "./02_Datasets/BIOGRID/")
download.file(url= "https://downloads.thebiogrid.org/Download/BioGRID/Release-Archive/BIOGRID-4.4.230/BIOGRID-ORGANISM-4.4.230.tab3.zip",
              destfile = "./02_Datasets/BIOGRID/BIOGRID-ORGANISM-4.4.230.tab3.zip") 
unzip(zipfile = "./02_Datasets/BIOGRID/BIOGRID-ORGANISM-4.4.230.tab3.zip", 
      exdir = "./02_Datasets/BIOGRID/BIOGRID-ORGANISM-4.4.230.tab3")
unlink(x = "./02_Datasets/BIOGRID/BIOGRID-ORGANISM-4.4.230.tab3.zip/", recursive = TRUE)

# Keep only Homo sapiens 
fn <- list.files(path = "./02_Datasets/BIOGRID/BIOGRID-ORGANISM-4.4.230.tab3/", full.names = TRUE)
fn_hs <- fn[str_detect(string = fn, pattern = "Homo_sapiens")]
fn_rm <- fn[str_detect(string = fn, pattern = "Homo_sapiens", negate = TRUE)]
unlink(x = fn_rm)
rm(fn, fn_rm)

# BIOGRID files to system variables 
Sys.setenv(BIOGRID_HS = fn_hs)
Sys.setenv(BIOGRID_HS_4R = paste0(fn_hs, ".4R", sep = ""))

```

```{bash eval=FALSE}

# Parse file to remove ' and # from file
sed "s/'//g" $BIOGRID_HS > $BIOGRID_HS_4R
sed "s/#//g" $BIOGRID_HS_4R > $BIOGRID_HS

```

```{r eval=FALSE}

# Remove intermediate file
unlink(x = paste0(fn_hs, ".4R", sep = ""))

```

```{r eval=FALSE}

# Homo sapiens interactions from BIOGRID
db <- read.table(file = fn_hs, sep = "\t", 
                 header = TRUE, stringsAsFactors = FALSE)
rm(fn_hs)

# Keep only physical interactions between Homo sapiens  
db <- db[which(db$Experimental.System.Type == "physical"), ]
db <- db[which(db$Organism.Name.Interactor.A == "Homo sapiens"), ]
db <- db[which(db$Organism.Name.Interactor.B == "Homo sapiens"), ]

# Interactions from Entrez to Ensembl Gene IDs
db <- data.frame(db, 
                 Ensembl_Gene_Interactor_A = entrez_to_ensembl$Ensembl_GeneID[match(x = db$Entrez.Gene.Interactor.A, table = entrez_to_ensembl$Entrez_ID)], 
                 Ensembl_Gene_Interactor_B = entrez_to_ensembl$Ensembl_GeneID[match(x = db$Entrez.Gene.Interactor.B, table = entrez_to_ensembl$Entrez_ID)])

# Remove interactions where either interactor cannot be mapped to Ensembl Gene IDs
nm <- unique(c(which(is.na(db$Ensembl_Gene_Interactor_A) == TRUE),
               which(is.na(db$Ensembl_Gene_Interactor_B) == TRUE)))
if(length(nm)> 0){ db <- db[-nm, ]}
rm(nm)

# Separate into interactions with a single and multiple Ensembl Gene IDs
db_m2m <- unique(c(which(str_detect(string = db$Ensembl_Gene_Interactor_A, pattern = "\\|") == TRUE),
                   which(str_detect(string = db$Ensembl_Gene_Interactor_B, pattern = "\\|") == TRUE)))

db_o2o <- setdiff(x = 1:nrow(db), y = db_m2m)
db_m2m <- db[db_m2m, ]
db_o2o <- db[db_o2o, ]

# Expand interactions with multiple ensembl gene IDs
output <- matrix(NA, nrow = 1, ncol = 2, dimnames = list("ToRemove", c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")))
for(i in c(1:nrow(db_m2m))){
  in_a <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_A[i], pattern = "\\|\\|"))
  in_b <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_B[i], pattern = "\\|\\|"))
  ab <- expand.grid("Ensembl_Gene_Interactor_A" = in_a, "Ensembl_Gene_Interactor_B" = in_b)
  output <- rbind(output, ab)
  
}
rm(in_a, in_b, ab, i)
output <- output[-1, ]

# Combine interactions that map a single and multiple Ensembl Gene IDs 
db_ens <- rbind(db_o2o[, c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")], 
                output)
rm(db, db_m2m, db_o2o, output)

# Remove self interactions
mark_self <- apply(db_ens, 1, FUN = function(x) length(unique(x)))
if(length(which(mark_self == 1))> 0){db_ens <- db_ens[-which(mark_self == 1), ]}
rm(mark_self)

# Re-order interactions to remove duplicates
db_ens_order <- t(apply(db_ens, 1, FUN = function(x) x[order(x)]))
db_ens <- unique(db_ens_order) 
rm(db_ens_order)

# Convert to data frame
rownames(db_ens) <- 1:nrow(db_ens)
db_ens <- as.data.frame(db_ens)

# Save as tab delimited file 
write.table(x = db_ens, 
            file = "./02_Datasets/BIOGRID/BIOGRID-ORGANISM-4.4.230.tab3_EnsemblGeneID.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

```

## BIOPLEX

```{r eval=FALSE}

# Data download
dir.create(path = "./02_Datasets/BIOPLEX/")


# Bioplex version 3.0 # 
# Same name different file size for HCT116
download.file(url = "https://bioplex.hms.harvard.edu/data/BioPlex_293T_Network_10K_Dec_2019.tsv",
                            destfile = "./02_Datasets/BIOPLEX/BioPlex_293T_Network_10K_Dec_2019.tsv")
download.file(url = "https://bioplex.hms.harvard.edu/data/BioPlex_HCT116_Network_5.5K_Dec_2019.tsv",
              destfile = "./02_Datasets/BIOPLEX/BioPlex_HCT116_Network_5.5K_Dec_2019.tsv")

```

### 293T

```{r eval=FALSE}

# Bioplex
db <- read.table(file = "./02_Datasets/BIOPLEX/BioPlex_293T_Network_10K_Dec_2019.tsv", sep = "\t", header = TRUE)

# Remove isoform info to mapp Uniprot ID
db$UniprotA <- sapply(str_split(string = db$UniprotA, pattern = "-"), "[", 1)
db$UniprotB <- sapply(str_split(string = db$UniprotB, pattern = "-"), "[", 1)

# Mapping through uniprot
db <- data.frame(db, 
                 Ensembl_Gene_Interactor_A = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$UniprotA, table = uniprot_to_ensembl$Uniprot_ID)],
                 Ensembl_Gene_Interactor_B = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$UniprotB, table = uniprot_to_ensembl$Uniprot_ID)])

# Remove interactions where either interactor cannot be mapped to Ensembl Gene IDs
nm <- unique(c(which(is.na(db$Ensembl_Gene_Interactor_A) == TRUE),
               which(is.na(db$Ensembl_Gene_Interactor_B) == TRUE)))
if(length(nm)> 0){ db <- db[-nm, ]}
rm(nm)

# Separate into interactions with a single and multiple Ensembl Gene IDs
db_m2m <- unique(c(which(str_detect(string = db$Ensembl_Gene_Interactor_A, pattern = "\\|") == TRUE),
                   which(str_detect(string = db$Ensembl_Gene_Interactor_B, pattern = "\\|") == TRUE)))

db_o2o <- setdiff(x = 1:nrow(db), y = db_m2m)
db_m2m <- db[db_m2m, ]
db_o2o <- db[db_o2o, ]

# Expand interactions with multiple ensembl gene IDs
output <- matrix(NA, nrow = 1, ncol = 2, dimnames = list("ToRemove", c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")))
for(i in c(1:nrow(db_m2m))){
  in_a <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_A[i], pattern = "\\|\\|"))
  in_b <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_B[i], pattern = "\\|\\|"))
  ab <- expand.grid("Ensembl_Gene_Interactor_A" = in_a, "Ensembl_Gene_Interactor_B" = in_b)
  output <- rbind(output, ab)
}
rm(in_a, in_b, ab, i)
output <- output[-1, ]

# Combine interactions that map a single and multiple Ensembl Gene IDs 
db_ens <- rbind(db_o2o[, c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")], 
                output)
rm(db, db_m2m, db_o2o, output)

# Remove self interactions
mark_self <- apply(db_ens, 1, FUN = function(x) length(unique(x)))
if(length(which(mark_self == 1))> 0){db_ens <- db_ens[-which(mark_self == 1), ]}
rm(mark_self)

# Re-order interactions to remove duplicates
db_ens_order <- t(apply(db_ens, 1, FUN = function(x) x[order(x)]))
db_ens <- unique(db_ens_order) 
rm(db_ens_order)

# Convert to data frame
rownames(db_ens) <- 1:nrow(db_ens)
db_ens <- as.data.frame(db_ens)

# Save as tab delimited file 
write.table(x = db_ens, 
            file = "./02_Datasets/BIOPLEX/BioPlex_293T_Network_10K_Dec_2019_EnsemblGeneID.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

```

### HCT116

```{r eval=FALSE}

# Bioplex
db <- read.table(file = "./02_Datasets/BIOPLEX/BioPlex_HCT116_Network_5.5K_Dec_2019.tsv", sep = "\t", header = TRUE)

# Remove isoform info to mapp Uniprot ID
db$UniprotA <- sapply(str_split(string = db$UniprotA, pattern = "-"), "[", 1)
db$UniprotB <- sapply(str_split(string = db$UniprotB, pattern = "-"), "[", 1)

# Mapping through uniprot
db <- data.frame(db, 
                 Ensembl_Gene_Interactor_A = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$UniprotA, table = uniprot_to_ensembl$Uniprot_ID)],
                 Ensembl_Gene_Interactor_B = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$UniprotB, table = uniprot_to_ensembl$Uniprot_ID)])

# Remove interactions where either interactor cannot be mapped to Ensembl Gene IDs
nm <- unique(c(which(is.na(db$Ensembl_Gene_Interactor_A) == TRUE),
               which(is.na(db$Ensembl_Gene_Interactor_B) == TRUE)))
if(length(nm)> 0){ db <- db[-nm, ]}
rm(nm)

# Separate into interactions with a single and multiple Ensembl Gene IDs
db_m2m <- unique(c(which(str_detect(string = db$Ensembl_Gene_Interactor_A, pattern = "\\|") == TRUE),
                   which(str_detect(string = db$Ensembl_Gene_Interactor_B, pattern = "\\|") == TRUE)))

db_o2o <- setdiff(x = 1:nrow(db), y = db_m2m)
db_m2m <- db[db_m2m, ]
db_o2o <- db[db_o2o, ]

# Expand interactions with multiple ensembl gene IDs
output <- matrix(NA, nrow = 1, ncol = 2, dimnames = list("ToRemove", c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")))
for(i in c(1:nrow(db_m2m))){
  in_a <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_A[i], pattern = "\\|\\|"))
  in_b <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_B[i], pattern = "\\|\\|"))
  ab <- expand.grid("Ensembl_Gene_Interactor_A" = in_a, "Ensembl_Gene_Interactor_B" = in_b)
  output <- rbind(output, ab)
}
rm(in_a, in_b, ab, i)
output <- output[-1, ]

# Combine interactions that map a single and multiple Ensembl Gene IDs 
db_ens <- rbind(db_o2o[, c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")], 
                output)
rm(db, db_m2m, db_o2o, output)

# Remove self interactions
mark_self <- apply(db_ens, 1, FUN = function(x) length(unique(x)))
if(length(which(mark_self == 1))> 0){db_ens <- db_ens[-which(mark_self == 1), ]}
rm(mark_self)

# Re-order interactions to remove duplicates
db_ens_order <- t(apply(db_ens, 1, FUN = function(x) x[order(x)]))
db_ens <- unique(db_ens_order) 
rm(db_ens_order)

# Convert to data frame
rownames(db_ens) <- 1:nrow(db_ens)
db_ens <- as.data.frame(db_ens)

# Save as tab delimited file 
write.table(x = db_ens, 
            file = "./02_Datasets/BIOPLEX/BioPlex_HCT116_Network_5.5K_Dec_2019_EnsemblGeneID.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

```

## CORUM

```{r eval=FALSE}

dir.create("./02_Datasets/CORUM/")

# Database is now (Feb 2024) online !

# All complexes
download.file("https://mips.helmholtz-muenchen.de/corum/download/releases/current/allComplexes.txt.zip", 
              destfile = "./02_Datasets/CORUM/allComplex.zip")

# Human complexes
download.file("https://mips.helmholtz-muenchen.de/corum/download/releases/current/humanComplexes.txt.zip", 
              destfile = "./02_Datasets/CORUM/humanComplexes.txt.zip")

# Partial complexes
download.file("https://mips.helmholtz-muenchen.de/corum/download/releases/current/partialComplexes.txt.zip", 
              destfile = "./02_Datasets/CORUM/partialComplexes.txt.zip")

# Unzip files
gunzip(zipfile = "./02_Datasets/CORUM/allComplexes.txt.zip", remove = TRUE)
gunzip(zipfile = "./02_Datasets/CORUM/humanComplexes.txt.zip", remove = TRUE)
gunzip(zipfile = "./02_Datasets/CORUM/partialComplexes.txt.zip", remove = TRUE)

Sys.setenv(CORUM_ALL = "./02_Datasets/CORUM/allComplexes.txt")
Sys.setenv(CORUM_ALL_TMP = "./02_Datasets/CORUM/allComplexes_tmp.txt")

Sys.setenv(CORUM_HUM = "./02_Datasets/CORUM/humanComplexes.txt")
Sys.setenv(CORUM_HUM_TMP = "./02_Datasets/CORUM/humanComplexes_tmp.txt")

Sys.setenv(CORUM_PAR = "./02_Datasets/CORUM/partialComplexes.txt")
Sys.setenv(CORUM_PAR_TMP = "./02_Datasets/CORUM/partialComplexes_tmp.txt")


```

```{bash eval=FALSE}

# Parse file to remove ' from file
sed "s/'//g" $CORUM_ALL > $CORUM_ALL_TMP
sed "s/'//g" $CORUM_HUM > $CORUM_HUM_TMP
sed "s/'//g" $CORUM_PAR > $CORUM_PAR_TMP

```

```{r eval=FALSE}

# Remove temporal file and rename
unlink(x = "./02_Datasets/CORUM/allComplexes.txt")
unlink(x = "./02_Datasets/CORUM/humanComplexes.txt")
unlink(x = "./02_Datasets/CORUM/partialComplexes.txt")

file.rename(from = "./02_Datasets/CORUM/allComplexes_tmp.txt", to = "./02_Datasets/CORUM/allComplexes.txt")
file.rename(from = "./02_Datasets/CORUM/humanComplexes_tmp.txt", to = "./02_Datasets/CORUM/humanComplexes.txt")
file.rename(from = "./02_Datasets/CORUM/partialComplexes_tmp.txt", to = "./02_Datasets/CORUM/partialComplexes.txt")

```

### All complexes

```{r eval=FALSE}

# CORUM
db <- read.table(file = "./02_Datasets/CORUM/allComplexes.txt", sep = "\t", header = TRUE)

# Rename column names 
colnames(db) <- str_replace_all(string = colnames(db), pattern = "\\.", replacement = "_")

# Reduce to human interactions
db <- db[which(db$Organism == "Human"), ]

# From complexes to pairs of interactions
output <- matrix(NA, nrow = 1, ncol = 2, 
                 dimnames = list("RemoveMe", c("Interactor_A", "Interactor_B")))

for(i in 1:nrow(db)){
  ci <- unlist(str_split(string = db$subunits_UniProt_IDs_[i], pattern = ";"))
  if(length(ci) > 0){
  output <- rbind(output, expand.grid("Interactor_A" = ci, "Interactor_B" = ci))  
  }
  rm(ci)
}
rm(i)
output <- output[-1, ]
rm(db)
db <- output
rm(output)
rownames(db) <- 1:nrow(db)

# Add ensembl annotations
db <- data.frame(db, 
                 Ensembl_Gene_Interactor_A = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$Interactor_A, table = uniprot_to_ensembl$Uniprot_ID)],
                 Ensembl_Gene_Interactor_B = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$Interactor_B, table = uniprot_to_ensembl$Uniprot_ID)])

# Remove interactions where either interactor cannot be mapped to Ensembl Gene IDs
nm <- unique(c(which(is.na(db$Ensembl_Gene_Interactor_A) == TRUE),
               which(is.na(db$Ensembl_Gene_Interactor_B) == TRUE)))
if(length(nm)> 0){ db <- db[-nm, ]}
rm(nm)

# Separate into interactions with a single and multiple Ensembl Gene IDs
db_m2m <- unique(c(which(str_detect(string = db$Ensembl_Gene_Interactor_A, pattern = "\\|") == TRUE),
                   which(str_detect(string = db$Ensembl_Gene_Interactor_B, pattern = "\\|") == TRUE)))

db_o2o <- setdiff(x = 1:nrow(db), y = db_m2m)
db_m2m <- db[db_m2m, ]
db_o2o <- db[db_o2o, ]

# Expand interactions with multiple ensembl gene IDs
output <- matrix(NA, nrow = 1, ncol = 2, dimnames = list("ToRemove", c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")))
for(i in c(1:nrow(db_m2m))){
  in_a <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_A[i], pattern = "\\|\\|"))
  in_b <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_B[i], pattern = "\\|\\|"))
  ab <- expand.grid("Ensembl_Gene_Interactor_A" = in_a, "Ensembl_Gene_Interactor_B" = in_b)
  output <- rbind(output, ab)
}
rm(in_a, in_b, ab, i)
output <- output[-1, ]

# Combine interactions that map a single and multiple Ensembl Gene IDs 
db_ens <- rbind(db_o2o[, c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")], 
                output)
rm(db, db_m2m, db_o2o, output)

# Remove self interactions
mark_self <- apply(db_ens, 1, FUN = function(x) length(unique(x)))
if(length(which(mark_self == 1))> 0){db_ens <- db_ens[-which(mark_self == 1), ]}
rm(mark_self)

# Re-order interactions to remove duplicates
db_ens_order <- t(apply(db_ens, 1, FUN = function(x) x[order(x)]))
db_ens <- unique(db_ens_order) 
rm(db_ens_order)

# Convert to data frame
rownames(db_ens) <- 1:nrow(db_ens)
db_ens <- as.data.frame(db_ens)

# Save as tab delimited file 
write.table(x = db_ens, 
            file = "./02_Datasets/CORUM/allComplexes_EnsemblGeneID.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

```

### Human complexes

```{r eval=FALSE}

# CORUM
db <- read.table(file = "./02_Datasets/CORUM/humanComplexes.txt", sep = "\t", header = TRUE)

# Rename column names 
colnames(db) <- str_replace_all(string = colnames(db), pattern = "\\.", replacement = "_")

# Reduce to human interactions
db <- db[which(db$Organism == "Human"), ]

# From complexes to pairs of interactions
output <- matrix(NA, nrow = 1, ncol = 2, dimnames = list("RemoveMe", c("Interactor_A", "Interactor_B")))
for(i in 1:nrow(db)){
  ci <- unlist(str_split(string = db$subunits_UniProt_IDs_[i], pattern = ";"))
  if(length(ci) > 0){
  output <- rbind(output, expand.grid("Interactor_A" = ci, "Interactor_B" = ci))  
  }
  rm(ci)
}
rm(i)
output <- output[-1, ]
rm(db)
db <- output
rm(output)
rownames(db) <- 1:nrow(db)

# Add ensembl annotations
db <- data.frame(db, 
                 Ensembl_Gene_Interactor_A = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$Interactor_A, table = uniprot_to_ensembl$Uniprot_ID)],
                 Ensembl_Gene_Interactor_B = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$Interactor_B, table = uniprot_to_ensembl$Uniprot_ID)])

# Remove interactions where either interactor cannot be mapped to Ensembl Gene IDs
nm <- unique(c(which(is.na(db$Ensembl_Gene_Interactor_A) == TRUE),
               which(is.na(db$Ensembl_Gene_Interactor_B) == TRUE)))
if(length(nm)> 0){ db <- db[-nm, ]}
rm(nm)

# Separate into interactions with a single and multiple Ensembl Gene IDs
db_m2m <- unique(c(which(str_detect(string = db$Ensembl_Gene_Interactor_A, pattern = "\\|") == TRUE),
                   which(str_detect(string = db$Ensembl_Gene_Interactor_B, pattern = "\\|") == TRUE)))

db_o2o <- setdiff(x = 1:nrow(db), y = db_m2m)
db_m2m <- db[db_m2m, ]
db_o2o <- db[db_o2o, ]

# Expand interactions with multiple ensembl gene IDs
output <- matrix(NA, nrow = 1, ncol = 2, dimnames = list("ToRemove", c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")))
for(i in c(1:nrow(db_m2m))){
  in_a <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_A[i], pattern = "\\|\\|"))
  in_b <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_B[i], pattern = "\\|\\|"))
  ab <- expand.grid("Ensembl_Gene_Interactor_A" = in_a, "Ensembl_Gene_Interactor_B" = in_b)
  output <- rbind(output, ab)
}
rm(in_a, in_b, ab, i)
output <- output[-1, ]

# Combine interactions that map a single and multiple Ensembl Gene IDs 
db_ens <- rbind(db_o2o[, c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")], 
                output)
rm(db, db_m2m, db_o2o, output)

# Remove self interactions
mark_self <- apply(db_ens, 1, FUN = function(x) length(unique(x)))
if(length(which(mark_self == 1))> 0){db_ens <- db_ens[-which(mark_self == 1), ]}
rm(mark_self)

# Re-order interactions to remove duplicates
db_ens_order <- t(apply(db_ens, 1, FUN = function(x) x[order(x)]))
db_ens <- unique(db_ens_order) 
rm(db_ens_order)

# Convert to data frame
rownames(db_ens) <- 1:nrow(db_ens)
db_ens <- as.data.frame(db_ens)

# Save as tab delimited file 
write.table(x = db_ens, 
            file = "./02_Datasets/CORUM/humanComplexes_EnsemblGeneID.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

```

### Partial complexes

```{r eval=FALSE}

# CORUM
db <- read.table(file = "./02_Datasets/CORUM/partialComplexes.txt", sep = "\t", header = TRUE)

# Rename column names 
colnames(db) <- str_replace_all(string = colnames(db), pattern = "\\.", replacement = "_")

# Reduce to human interactions
db <- db[which(db$Organism == "Human"), ]

# From complexes to pairs of interactions
output <- matrix(NA, nrow = 1, ncol = 2, dimnames = list("RemoveMe", c("Interactor_A", "Interactor_B")))
for(i in 1:nrow(db)){
  ci <- unlist(str_split(string = db$subunits_UniProt_IDs_[i], pattern = ";"))
  if(length(ci) > 0){
  output <- rbind(output, expand.grid("Interactor_A" = ci, "Interactor_B" = ci))  
  }
  rm(ci)
}
rm(i)
output <- output[-1, ]
rm(db)
db <- output
rm(output)
rownames(db) <- 1:nrow(db)

# Add ensembl annotations
db <- data.frame(db, 
                 Ensembl_Gene_Interactor_A = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$Interactor_A, table = uniprot_to_ensembl$Uniprot_ID)],
                 Ensembl_Gene_Interactor_B = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$Interactor_B, table = uniprot_to_ensembl$Uniprot_ID)])

# Remove interactions where either interactor cannot be mapped to Ensembl Gene IDs
nm <- unique(c(which(is.na(db$Ensembl_Gene_Interactor_A) == TRUE),
               which(is.na(db$Ensembl_Gene_Interactor_B) == TRUE)))
if(length(nm)> 0){ db <- db[-nm, ]}
rm(nm)

# Separate into interactions with a single and multiple Ensembl Gene IDs
db_m2m <- unique(c(which(str_detect(string = db$Ensembl_Gene_Interactor_A, pattern = "\\|") == TRUE),
                   which(str_detect(string = db$Ensembl_Gene_Interactor_B, pattern = "\\|") == TRUE)))

db_o2o <- setdiff(x = 1:nrow(db), y = db_m2m)
db_m2m <- db[db_m2m, ]
db_o2o <- db[db_o2o, ]

# Expand interactions with multiple ensembl gene IDs
output <- matrix(NA, nrow = 1, ncol = 2, dimnames = list("ToRemove", c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")))
for(i in c(1:nrow(db_m2m))){
  in_a <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_A[i], pattern = "\\|\\|"))
  in_b <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_B[i], pattern = "\\|\\|"))
  ab <- expand.grid("Ensembl_Gene_Interactor_A" = in_a, "Ensembl_Gene_Interactor_B" = in_b)
  output <- rbind(output, ab)
}
rm(in_a, in_b, ab, i)
output <- output[-1, ]

# Combine interactions that map a single and multiple Ensembl Gene IDs 
db_ens <- rbind(db_o2o[, c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")], 
                output)
rm(db, db_m2m, db_o2o, output)

# Remove self interactions
mark_self <- apply(db_ens, 1, FUN = function(x) length(unique(x)))
if(length(which(mark_self == 1))> 0){db_ens <- db_ens[-which(mark_self == 1), ]}
rm(mark_self)

# Re-order interactions to remove duplicates
db_ens_order <- t(apply(db_ens, 1, FUN = function(x) x[order(x)]))
db_ens <- unique(db_ens_order) 
rm(db_ens_order)

# Convert to data frame
rownames(db_ens) <- 1:nrow(db_ens)
db_ens <- as.data.frame(db_ens)

# Save as tab delimited file 
write.table(x = db_ens, 
            file = "./02_Datasets/CORUM/partialComplexes_EnsemblGeneID.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

```


## HITPREDICT

```{bash eval=FALSE}

# Download interactions - # August 2023 last update.
# You require to write an email to ashwini@combinatics.com to request access 
# Just send name, affiliation and brief description of use to get access link.
# tar zxvf ./02_Datasets/HITPREDICT/H_sapiens_interactions.txt.tgz
# tar zxvf ./02_Datasets/HITPREDICT/HitPredict_interactions.txt.tgz

```

```{r eval=FALSE}

# HitPredict latest
db <- read.table(file = "./02_Datasets/HITPREDICT/H_sapiens_interactions.txt", sep = "\t", header = TRUE, skip = 4)

# Mapping through uniprot
db <- data.frame(db, 
                 Ensembl_Gene_Interactor_A = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$Uniprot1, table = uniprot_to_ensembl$Uniprot_ID)],
                 Ensembl_Gene_Interactor_B = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$Uniprot2, table = uniprot_to_ensembl$Uniprot_ID)])

# Remove interactions where either interactor cannot be mapped to Ensembl Gene IDs
nm <- unique(c(which(is.na(db$Ensembl_Gene_Interactor_A) == TRUE),
               which(is.na(db$Ensembl_Gene_Interactor_B) == TRUE)))
if(length(nm)> 0){ db <- db[-nm, ]}
rm(nm)

# Separate into interactions with a single and multiple Ensembl Gene IDs
db_m2m <- unique(c(which(str_detect(string = db$Ensembl_Gene_Interactor_A, pattern = "\\|") == TRUE),
                   which(str_detect(string = db$Ensembl_Gene_Interactor_B, pattern = "\\|") == TRUE)))

db_o2o <- setdiff(x = 1:nrow(db), y = db_m2m)
db_m2m <- db[db_m2m, ]
db_o2o <- db[db_o2o, ]

# Expand interactions with multiple ensembl gene IDs
output <- matrix(NA, nrow = 1, ncol = 2, dimnames = list("ToRemove", c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")))
for(i in c(1:nrow(db_m2m))){
  in_a <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_A[i], pattern = "\\|\\|"))
  in_b <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_B[i], pattern = "\\|\\|"))
  ab <- expand.grid("Ensembl_Gene_Interactor_A" = in_a, "Ensembl_Gene_Interactor_B" = in_b)
  output <- rbind(output, ab)
}
rm(in_a, in_b, ab, i)
output <- output[-1, ]

# Combine interactions that map a single and multiple Ensembl Gene IDs 
db_ens <- rbind(db_o2o[, c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")], 
                output)
rm(db, db_m2m, db_o2o, output)

# Remove self interactions
mark_self <- apply(db_ens, 1, FUN = function(x) length(unique(x)))
if(length(which(mark_self == 1))> 0){db_ens <- db_ens[-which(mark_self == 1), ]}
rm(mark_self)

# Re-order interactions to remove duplicates
db_ens_order <- t(apply(db_ens, 1, FUN = function(x) x[order(x)]))
db_ens <- unique(db_ens_order) 
rm(db_ens_order)

# Convert to data frame
rownames(db_ens) <- 1:nrow(db_ens)
db_ens <- as.data.frame(db_ens)

# Save as tab delimited file 
write.table(x = db_ens, 
            file = "./02_Datasets/HITPREDICT/H_sapiens_interactions_EnsemblGeneID.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

```


## HURI

```{r eval=FALSE}

dir.create(path = "./02_Datasets/HURI/")
download.file(url = "http://www.interactome-atlas.org/data/HuRI.tsv",
              destfile = "./02_Datasets/HURI/HuRI.tsv")

download.file(url = "http://www.interactome-atlas.org/data/HI-union.tsv", 
              destfile = "./02_Datasets/HURI/HI-union.tsv")

download.file(url = "http://www.interactome-atlas.org/data/Test_space_screens-19.tsv", 
              destfile = "./02_Datasets/HURI/Test_space_screens-19.tsv")

download.file(url = "http://www.interactome-atlas.org/data/Lit-BM.tsv", 
              destfile = "./02_Datasets/HURI/Lit-BM.tsv")

download.file(url = "http://www.interactome-atlas.org/data/Yang-16.tsv", 
              destfile = "./02_Datasets/HURI/Yang-16.tsv")

download.file(url = "http://www.interactome-atlas.org/data/HI-II-14.tsv", 
              destfile = "./02_Datasets/HURI/HI-II-14.tsv")

download.file(url = "http://www.interactome-atlas.org/data/Venkatesan-09.tsv", 
              destfile = "./02_Datasets/HURI/Venkatesan-09.tsv")

download.file(url = "http://www.interactome-atlas.org/data/H-I-05.tsv",
              destfile = "./02_Datasets/HURI/H-I-05.tsv")

download.file(url = "http://www.interactome-atlas.org/data/Yu-11.tsv",
              destfile = "./02_Datasets/HURI/Yu-11.tsv")

fn <- list.files(path = "./02_Datasets/HURI/", full.names = TRUE)

```

```{r eval=FALSE}

for (i in c(1:length(fn))){
  
  # HuRI
  db_ens <- read.table(file = fn[i], sep = "\t", header = FALSE)

  # Remove self interactions
  mark_self <- apply(db_ens, 1, FUN = function(x) length(unique(x)))
  if(length(which(mark_self == 1))> 0){db_ens <- db_ens[-which(mark_self == 1), ]}
  rm(mark_self)

  # Re-order interactions to remove duplicates
  db_ens_order <- t(apply(db_ens, 1, FUN = function(x) x[order(x)]))
  db_ens <- unique(db_ens_order) 
  rm(db_ens_order)

  out_name <- paste0(str_replace(string = fn[i], 
                                 pattern = ".tsv", 
                                 replacement = ""), 
                     "_EnsemblGeneID.txt", collapse = "")

  # Save as tab delimited file 
  write.table(x = db_ens, 
              file = out_name, 
              sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)
  rm(db_ens, out_name)
}
rm(fn, i)

```


## INTACT

```{r eval=FALSE}

# Data download
dir.create(path = "./02_Datasets/INTACT/")
download.file(url = "https://ftp.ebi.ac.uk/pub/databases/intact/current/psimitab/intact.zip",
              destfile = "./02_Datasets/INTACT/intact.zip") 
unzip(zipfile = "./02_Datasets/INTACT/intact.zip", exdir = "./02_Datasets/INTACT/")
unlink(x = "./02_Datasets/INTACT/intact.zip")

# INTACT files to system variables 
Sys.setenv(INTACT = "./02_Datasets/INTACT/intact.txt")
Sys.setenv(INTACT_TMP = "./02_Datasets/INTACT/intact_tmp.txt")

```

```{bash eval=FALSE}

# Parse file to remove ' and # from file
sed "s/'//g" $INTACT > $INTACT_TMP
sed "s/#//g" $INTACT_TMP > $INTACT

```

```{r eval=FALSE}

# Remove intermediate files
unlink(x = "./02_Datasets/INTACT/intact_tmp.txt")

```

```{r eval=FALSE}

# Intact db
db <- read.table(file = "./02_Datasets/INTACT/intact.txt", sep = "\t", quote = "", header = TRUE)

# Replace "." from colnames 
colnames(db) <- str_replace_all(string = colnames(db), pattern = "\\.\\.", replacement = ".")
colnames(db) <- str_replace_all(string = colnames(db), pattern = "\\.", replacement = "_")

# Only within human interactions
hsap <- intersect(which(str_detect(string = db$Taxid_interactor_A, pattern = "taxid:9606") == TRUE), 
                  which(str_detect(string = db$Taxid_interactor_B, pattern = "taxid:9606") == TRUE))
db <- db[hsap, ]
rm(hsap)

# Mapping through uniprot
uni <- intersect(which(str_detect(string = db$ID_s_interactor_A, pattern = "uniprotkb") == TRUE),
                 which(str_detect(string = db$ID_s_interactor_B, pattern = "uniprotkb") == TRUE))
db <- db[uni, ]
rm(uni)
db$ID_s_interactor_A <- str_replace(string = db$ID_s_interactor_A, pattern = "uniprotkb:", replacement = "")
db$ID_s_interactor_B <- str_replace(string = db$ID_s_interactor_B, pattern = "uniprotkb:", replacement = "")

db <- data.frame(db, 
                 Ensembl_Gene_Interactor_A = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$ID_s_interactor_A, table = uniprot_to_ensembl$Uniprot_ID)],
                 Ensembl_Gene_Interactor_B = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$ID_s_interactor_B, table = uniprot_to_ensembl$Uniprot_ID)])

# Remove interactions where either interactor cannot be mapped to Ensembl Gene IDs
nm <- unique(c(which(is.na(db$Ensembl_Gene_Interactor_A) == TRUE),
               which(is.na(db$Ensembl_Gene_Interactor_B) == TRUE)))
if(length(nm)> 0){ db <- db[-nm, ]}
rm(nm)

# Separate into interactions with a single and multiple Ensembl Gene IDs
db_m2m <- unique(c(which(str_detect(string = db$Ensembl_Gene_Interactor_A, pattern = "\\|") == TRUE),
                   which(str_detect(string = db$Ensembl_Gene_Interactor_B, pattern = "\\|") == TRUE)))

db_o2o <- setdiff(x = 1:nrow(db), y = db_m2m)
db_m2m <- db[db_m2m, ]
db_o2o <- db[db_o2o, ]

# Expand interactions with multiple ensembl gene IDs
output <- matrix(NA, nrow = 1, ncol = 2, dimnames = list("ToRemove", c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")))
for(i in c(1:nrow(db_m2m))){
  in_a <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_A[i], pattern = "\\|\\|"))
  in_b <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_B[i], pattern = "\\|\\|"))
  ab <- expand.grid("Ensembl_Gene_Interactor_A" = in_a, "Ensembl_Gene_Interactor_B" = in_b)
  output <- rbind(output, ab)
}
rm(in_a, in_b, ab, i)
output <- output[-1, ]

# Combine interactions that map a single and multiple Ensembl Gene IDs 
db_ens <- rbind(db_o2o[, c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")], 
                output)
rm(db, db_m2m, db_o2o, output)

# Remove self interactions
mark_self <- apply(db_ens, 1, FUN = function(x) length(unique(x)))
if(length(which(mark_self == 1))> 0){db_ens <- db_ens[-which(mark_self == 1), ]}
rm(mark_self)

# Re-order interactions to remove duplicates
db_ens_order <- t(apply(db_ens, 1, FUN = function(x) x[order(x)]))
db_ens <- unique(db_ens_order) 
rm(db_ens_order)

# Convert to data frame
rownames(db_ens) <- 1:nrow(db_ens)
db_ens <- as.data.frame(db_ens)

# Save as tab delimited file 
write.table(x = db_ens, 
            file = "./02_Datasets/INTACT/intact_EnsemblGeneID.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

```

## MINT

```{r eval=FALSE}

# Data download
dir.create(path = "./02_Datasets/MINT/")

download.file(url = "http://www.ebi.ac.uk/Tools/webservices/psicquic/mint/webservices/current/search/query/species:human",
              destfile = "./02_Datasets/MINT/species_human.txt")

Sys.setenv(MINT = "./02_Datasets/MINT/species_human.txt")
Sys.setenv(MINT_4R = "./02_Datasets/MINT/species_human_tmp.txt")

```

```{bash eval=FALSE}

# Parse file to remove ' and # from file
sed "s/'//g" $MINT > $MINT_4R
sed "s/#//g" $MINT_4R > $MINT

```

```{r eval=FALSE}

# Remove temporal file
unlink(x = "./02_Datasets/MINT/species_human_tmp.txt")

```

```{r eval=FALSE}

# Mint
db <- read.delim("./02_Datasets/MINT/species_human.txt", sep = "\t", quote = NULL, header = FALSE, stringsAsFactors = FALSE)

# Rename a few columns
colnames(db)[1:2] <- c("Interactor_A", "Interactor_B")
colnames(db)[10:11] <- c("Taxid_interactor_A", "Taxid_interactor_B")

# Only within human interactions
hsap <- intersect(which(str_detect(string = db$Taxid_interactor_A, pattern = "taxid:9606") == TRUE), 
                  which(str_detect(string = db$Taxid_interactor_B, pattern = "taxid:9606") == TRUE))
db <- db[hsap, ]
rm(hsap)

# Mapping through uniprot
uni <- intersect(which(str_detect(string = db$Interactor_A, pattern = "uniprotkb") == TRUE),
                 which(str_detect(string = db$Interactor_B, pattern = "uniprotkb") == TRUE))
db <- db[uni, ]
rm(uni)
db$Interactor_A <- str_replace(string = db$Interactor_A, pattern = "uniprotkb:", replacement = "")
db$Interactor_B <- str_replace(string = db$Interactor_B, pattern = "uniprotkb:", replacement = "")

# Mapping through uniprot
db <- data.frame(db, 
                 Ensembl_Gene_Interactor_A = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$Interactor_A, table = uniprot_to_ensembl$Uniprot_ID)],
                 Ensembl_Gene_Interactor_B = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$Interactor_B, table = uniprot_to_ensembl$Uniprot_ID)])

# Remove interactions where either interactor cannot be mapped to Ensembl Gene IDs
nm <- unique(c(which(is.na(db$Ensembl_Gene_Interactor_A) == TRUE),
               which(is.na(db$Ensembl_Gene_Interactor_B) == TRUE)))
if(length(nm)> 0){ db <- db[-nm, ]}
rm(nm)

# Separate into interactions with a single and multiple Ensembl Gene IDs
db_m2m <- unique(c(which(str_detect(string = db$Ensembl_Gene_Interactor_A, pattern = "\\|") == TRUE),
                   which(str_detect(string = db$Ensembl_Gene_Interactor_B, pattern = "\\|") == TRUE)))

db_o2o <- setdiff(x = 1:nrow(db), y = db_m2m)
db_m2m <- db[db_m2m, ]
db_o2o <- db[db_o2o, ]

# Expand interactions with multiple ensembl gene IDs
output <- matrix(NA, nrow = 1, ncol = 2, dimnames = list("ToRemove", c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")))
for(i in c(1:nrow(db_m2m))){
  in_a <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_A[i], pattern = "\\|\\|"))
  in_b <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_B[i], pattern = "\\|\\|"))
  ab <- expand.grid("Ensembl_Gene_Interactor_A" = in_a, "Ensembl_Gene_Interactor_B" = in_b)
  output <- rbind(output, ab)
}
rm(in_a, in_b, ab, i)
output <- output[-1, ]

# Combine interactions that map a single and multiple Ensembl Gene IDs 
db_ens <- rbind(db_o2o[, c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")], 
                output)
rm(db, db_m2m, db_o2o, output)

# Remove self interactions
mark_self <- apply(db_ens, 1, FUN = function(x) length(unique(x)))
if(length(which(mark_self == 1))> 0){db_ens <- db_ens[-which(mark_self == 1), ]}
rm(mark_self)

# Re-order interactions to remove duplicates
db_ens_order <- t(apply(db_ens, 1, FUN = function(x) x[order(x)]))
db_ens <- unique(db_ens_order) 
rm(db_ens_order)

# Convert to data frame
rownames(db_ens) <- 1:nrow(db_ens)
db_ens <- as.data.frame(db_ens)

# Save as tab delimited file 
write.table(x = db_ens, 
            file = "./02_Datasets/MINT/species_human_EnsemblGeneID.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

```

## REACTOME

```{r eval=FALSE}

# Data download
dir.create(path = "./02_Datasets/REACTOME/")
download.file(url = "https://reactome.org/download/current/interactors/reactome.homo_sapiens.interactions.tab-delimited.txt",
              destfile = "./02_Datasets/REACTOME/reactome.homo_sapiens.interactions.tab-delimited.txt") 


# REACTOME files to system variables 
Sys.setenv(REACTOME = "./02_Datasets/REACTOME/reactome.homo_sapiens.interactions.tab-delimited.txt")
Sys.setenv(REACTOME_TMP = "./02_Datasets/REACTOME/reactome.homo_sapiens.interactions.tab-delimited_tmp.txt")

```

```{bash eval=FALSE}

# Parse file to remove ' and # from file
sed "s/'//g" $REACTOME > $REACTOME_TMP
sed "s/#//g" $REACTOME_TMP > $REACTOME

```

```{r eval=FALSE}

# Remove intermediate files
unlink(x = "./02_Datasets/REACTOME/reactome.homo_sapiens.interactions.tab-delimited_tmp.txt")

```

```{r eval=FALSE}

# Reactome
db <- read.table(file = "./02_Datasets/REACTOME/reactome.homo_sapiens.interactions.tab-delimited.txt", sep = "\t", quote = "", header = TRUE)

# Replace "." from colnames 
colnames(db) <- str_replace_all(string = colnames(db), pattern = "\\.", replacement = "_")

# Mapping through uniprot
uni <- intersect(which(str_detect(string = db$Interactor_1_uniprot_id, pattern = "uniprotkb") == TRUE),
                 which(str_detect(string = db$Interactor_2_uniprot_id, pattern = "uniprotkb") == TRUE))
db <- db[uni, ]
rm(uni)
db$Interactor_1_uniprot_id <- str_replace(string = db$Interactor_1_uniprot_id, pattern = "uniprotkb:", replacement = "")
db$Interactor_2_uniprot_id <- str_replace(string = db$Interactor_2_uniprot_id, pattern = "uniprotkb:", replacement = "")

db <- data.frame(db, 
                 Ensembl_Gene_Interactor_A = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$Interactor_1_uniprot_id, table = uniprot_to_ensembl$Uniprot_ID)],
                 Ensembl_Gene_Interactor_B = uniprot_to_ensembl$Ensembl_GeneID[match(x = db$Interactor_2_uniprot_id, table = uniprot_to_ensembl$Uniprot_ID)])

# Remove interactions where either interactor cannot be mapped to Ensembl Gene IDs
nm <- unique(c(which(is.na(db$Ensembl_Gene_Interactor_A) == TRUE),
               which(is.na(db$Ensembl_Gene_Interactor_B) == TRUE)))
if(length(nm)> 0){ db <- db[-nm, ]}
rm(nm)

# Separate into interactions with a single and multiple Ensembl Gene IDs
db_m2m <- unique(c(which(str_detect(string = db$Ensembl_Gene_Interactor_A, pattern = "\\|") == TRUE),
                   which(str_detect(string = db$Ensembl_Gene_Interactor_B, pattern = "\\|") == TRUE)))

db_o2o <- setdiff(x = 1:nrow(db), y = db_m2m)
db_m2m <- db[db_m2m, ]
db_o2o <- db[db_o2o, ]

# Expand interactions with multiple ensembl gene IDs
output <- matrix(NA, nrow = 1, ncol = 2, dimnames = list("ToRemove", c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")))
for(i in c(1:nrow(db_m2m))){
  in_a <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_A[i], pattern = "\\|\\|"))
  in_b <- unlist(str_split(string = db_m2m$Ensembl_Gene_Interactor_B[i], pattern = "\\|\\|"))
  ab <- expand.grid("Ensembl_Gene_Interactor_A" = in_a, "Ensembl_Gene_Interactor_B" = in_b)
  output <- rbind(output, ab)
}
rm(in_a, in_b, ab, i)
output <- output[-1, ]

# Combine interactions that map a single and multiple Ensembl Gene IDs 
db_ens <- rbind(db_o2o[, c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")], 
                output)
rm(db, db_m2m, db_o2o, output)

# Remove self interactions
mark_self <- apply(db_ens, 1, FUN = function(x) length(unique(x)))
if(length(which(mark_self == 1))> 0){db_ens <- db_ens[-which(mark_self == 1), ]}
rm(mark_self)

# Re-order interactions to remove duplicates
db_ens_order <- t(apply(db_ens, 1, FUN = function(x) x[order(x)]))
db_ens <- unique(db_ens_order) 
rm(db_ens_order)

# Convert to data frame
rownames(db_ens) <- 1:nrow(db_ens)
db_ens <- as.data.frame(db_ens)

# Save as tab delimited file 
write.table(x = db_ens, 
            file = "./02_Datasets/REACTOME/reactome.homo_sapiens.interactions.tab-delimited_EnsemblGeneID.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

```

## STRING

```{r eval=FALSE}

# Data download
dir.create(path = "./02_Datasets/STRING/")

download.file(url = "https://stringdb-downloads.org/download/protein.links.v12.0/9606.protein.links.v12.0.txt.gz",
              destfile = "./02_Datasets/STRING/9606.protein.links.v12.0.txt.gz")

download.file(url = "https://stringdb-downloads.org/download/protein.physical.links.v12.0/9606.protein.physical.links.v12.0.txt.gz",
              destfile = "./02_Datasets/STRING/9606.protein.physical.links.v12.0.txt.gz")

gunzip("./02_Datasets/STRING/9606.protein.links.v12.0.txt.gz", remove = TRUE)
gunzip("./02_Datasets/STRING/9606.protein.physical.links.v12.0.txt.gz", remove = TRUE)

```

### Protein links

```{r eval=FALSE}

# String
db <- read.table(file = "./02_Datasets/STRING/9606.protein.links.v12.0.txt", header = TRUE)
db$protein1 <- str_replace(string = db$protein1, pattern = "9606.", replacement = "")
db$protein2 <- str_replace(string = db$protein2, pattern = "9606.", replacement = "")

# Mapping through protein ID
db <- data.frame(db, 
                 Ensembl_Gene_Interactor_A = proteinID_to_ensembl$Ensembl_GeneID[match(x = db$protein1, table = proteinID_to_ensembl$Protein_ID)],
                 Ensembl_Gene_Interactor_B = proteinID_to_ensembl$Ensembl_GeneID[match(x = db$protein2, table = proteinID_to_ensembl$Protein_ID)])

# Remove interactions where either interactor cannot be mapped to Ensembl Gene IDs
nm <- unique(c(which(is.na(db$Ensembl_Gene_Interactor_A) == TRUE),
               which(is.na(db$Ensembl_Gene_Interactor_B) == TRUE)))
if(length(nm)> 0){ db <- db[-nm, ]}
rm(nm)

# Check if there are multiple Ensembl Gene IDs
#db_m2m <- unique(c(which(str_detect(string = db$Ensembl_Gene_Interactor_A, pattern = "\\|") == TRUE),
#                   which(str_detect(string = db$Ensembl_Gene_Interactor_B, pattern = "\\|") == TRUE)))
#length(db_m2m)

# Combine interactions that map a single and multiple Ensembl Gene IDs 
db_ens <- db[, c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")]
rm(db)

# Remove self interactions
mark_self <- apply(db_ens, 1, FUN = function(x) length(unique(x)))
if(length(which(mark_self == 1))> 0){db_ens <- db_ens[-which(mark_self == 1), ]}
rm(mark_self)

# Re-order interactions to remove duplicates
db_ens_order <- t(apply(db_ens, 1, FUN = function(x) x[order(x)]))
db_ens <- unique(db_ens_order) 
rm(db_ens_order)

# Convert to data frame
rownames(db_ens) <- 1:nrow(db_ens)
db_ens <- as.data.frame(db_ens)

# Save as tab delimited file 
write.table(x = db_ens, 
            file = "./02_Datasets/STRING/9606.protein.links.v12.0_EnsemblGeneID.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

```

### Protein physical

```{r eval=FALSE}

# String
db <- read.table(file = "./02_Datasets/STRING/9606.protein.physical.links.v12.0.txt", header = TRUE)
db$protein1 <- str_replace(string = db$protein1, pattern = "9606.", replacement = "")
db$protein2 <- str_replace(string = db$protein2, pattern = "9606.", replacement = "")

# Mapping through protein ID
db <- data.frame(db, 
                 Ensembl_Gene_Interactor_A = proteinID_to_ensembl$Ensembl_GeneID[match(x = db$protein1, table = proteinID_to_ensembl$Protein_ID)],
                 Ensembl_Gene_Interactor_B = proteinID_to_ensembl$Ensembl_GeneID[match(x = db$protein2, table = proteinID_to_ensembl$Protein_ID)])

# Remove interactions where either interactor cannot be mapped to Ensembl Gene IDs
nm <- unique(c(which(is.na(db$Ensembl_Gene_Interactor_A) == TRUE),
               which(is.na(db$Ensembl_Gene_Interactor_B) == TRUE)))
if(length(nm)> 0){ db <- db[-nm, ]}
rm(nm)

# Check if there are multiple Ensembl Gene IDs
#db_m2m <- unique(c(which(str_detect(string = db$Ensembl_Gene_Interactor_A, pattern = "\\|") == TRUE),
#                   which(str_detect(string = db$Ensembl_Gene_Interactor_B, pattern = "\\|") == TRUE)))
#length(db_m2m)

# Combine interactions that map a single and multiple Ensembl Gene IDs 
db_ens <- db[, c("Ensembl_Gene_Interactor_A", "Ensembl_Gene_Interactor_B")]
rm(db)

# Remove self interactions
mark_self <- apply(db_ens, 1, FUN = function(x) length(unique(x)))
if(length(which(mark_self == 1))> 0){db_ens <- db_ens[-which(mark_self == 1), ]}
rm(mark_self)

# Re-order interactions to remove duplicates
db_ens_order <- t(apply(db_ens, 1, FUN = function(x) x[order(x)]))
db_ens <- unique(db_ens_order) 
rm(db_ens_order)

# Convert to data frame
rownames(db_ens) <- 1:nrow(db_ens)
db_ens <- as.data.frame(db_ens)

# Save as tab delimited file 
write.table(x = db_ens, 
            file = "./02_Datasets/STRING/9606.protein.physical.links.v12.0_EnsemblGeneID.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

```

## Overview resources

```{r}

# List all protein protein interactions given in Ensembl Gene IDs
fn <- list.files(path = "./02_Datasets/", recursive = TRUE, pattern = "EnsemblGeneID", full.names = TRUE)

gfn <- sapply(str_split(string = fn, pattern = "\\/"), "[", 3)
gfd <- str_replace(string = sapply(str_split(string = fn, pattern = "\\/"), "[", 4), 
                   pattern = "_EnsemblGeneID.txt", replacement = "")

# Relabel datasets # 
names_datasets <- data.frame(fn_sub = c("9606_noISI_Q3", 
                                        "BIOGRID-ORGANISM-Homo_sapiens-4.4.230.tab3",
                                        "BioPlex_293T_Network_10K_Dec_2019", "BioPlex_HCT116_Network_5.5K_Dec_2019",
                                        "allComplexes", "humanComplexes", "partialComplexes", 
                                        "H_sapiens_interactions",
                                        "H-I-05", "HI-II-14", "HI-union", "HuRI", "Lit-BM", 
                                        "Test_space_screens-19", "Venkatesan-09", "Yang-16", "Yu-11",
                                        "intact", 
                                        "species_human",
                                        "reactome.homo_sapiens.interactions.tab-delimited",
                                        "9606.protein.links.v12.0", "9606.protein.physical.links.v12.0"),
                             Dataset = c("APID", 
                                         "BIOGRID", 
                                         "BIOPLEX_293T", "BIOPLEX_HCT116",
                                         "CORUM_All", "CORUM_human", "CORUM_partial",  
                                         "HITPREDICT_latest",
                                         "HURI_H-I-05", "HURI_HI-II-14", "HURI_HI-union", "HURI_HuRI", "HURI_Lit-BM",
                                         "HURI_Test19", "HURI_Venkatesan", "HURI_Yang", "HURI_Yu",
                                         "INTACT", 
                                         "MINT", 
                                         "REACTOME",
                                         "STRING_Links", "STRING_Physical"))

# Obtain an overview of each
output <- matrix(NA, length(fn), 4, 
                 dimnames = list(gfd, c("Genes", "Interactions", "Mean per gene", "Median per gene")))

for (i in c(1:length(fn))){ 
  fni <- read.table(file = fn[i], sep = "\t", stringsAsFactors = FALSE)
  output[i, 1] <- length(unique(c(fni$V1, fni$V2)))
  output[i, 2] <- nrow(fni)
  output[i, 3] <- mean(table(c(fni$V1, fni$V2)))
  output[i, 4] <- median(table(c(fni$V1, fni$V2)))
}
rm(fni, i)

output <- data.frame(Source = gfn, fn_sub = gfd, output)
output <- merge(output, names_datasets, by = "fn_sub")

```

```{r fig.height=4, fig.width=5.5}

# Number of genes
g1 <- ggplot(data = output, mapping = aes(x = Dataset, y = Genes, fill = Source, label = Genes)) + 
  geom_col() + geom_text(angle = 90, size = 3) + theme_classic() +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0))

# Number of interactions
g2 <- ggplot(data = output, mapping = aes(x = Dataset, y = Interactions, 
                                          fill = Source, label = Interactions)) + 
  geom_col() + geom_text(angle = 90, size = 3) + theme_classic() +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0))

# Mean interactions per gene
g3 <- ggplot(data = output, mapping = aes(x = Dataset, y = Mean.per.gene, 
                                          fill = Source)) + 
  geom_col() + theme_classic() +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0))

# Median interactions per gene
g4 <- ggplot(data = output, mapping = aes(x = Dataset, y = Median.per.gene, 
                                          fill = Source)) + 
  geom_col() + theme_classic() +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0))

# Overview
g5 <- ggplot(data = output, mapping = aes(x = log10(Interactions), 
                                          y = Genes, 
                                          color = Source, 
                                          size = Median.per.gene,
                                          label = Dataset)) + 
  geom_point() + 
  geom_text_repel() + 
  theme_classic() + 
  scale_color_met_d(name = "Signac")

g1
g2
g3
g4
g5

rm(g1, g2, g3, g4, g5, output)

```

## Merged PPI network

```{r eval=FALSE}

# Create directory for merged PPI
dir.create(path = "./03_PPIMerged/")

```

### All

```{r}

# Combine all datasets
PPIall <- read.table(fn[1], header = FALSE)

for(i in c(2:length(fn))){
  ppi <- read.table(file = fn[i], header = FALSE)
  PPIall <- rbind(PPIall, ppi)
  rm(ppi)
}

# Keep only unique interactions
PPIall <- unique(PPIall)

fn_ppi <- paste0(c("./03_PPIMerged/",
                   "MergedPPI_",
                   paste0(sapply(str_split(date(), " "), "[", c(5, 2)), collapse = ""),
                   # unique(gfn), # Optional to include names of datasets in output file
                   ".txt"),
                 collapse = "")

write.table(x = PPIall, file = fn_ppi, sep = "\t", 
            quote = FALSE, row.names = FALSE, col.names = FALSE)

rm(PPIall)

```

```{r}

# Combine datasets except string links
fn <- setdiff(fn, y = "./02_Datasets/STRING/9606.protein.links.v12.0_EnsemblGeneID.txt")

PPIall <- read.table(fn[1], header = FALSE)

for(i in c(2:length(fn))){
  ppi <- read.table(file = fn[i], header = FALSE)
  PPIall <- rbind(PPIall, ppi)
  rm(ppi)
}

# Keep only unique interactions
PPIall <- unique(PPIall)

fn_ppi <- paste0(c("./03_PPIMerged/",
                   "MergedPPI_",
                   paste0(sapply(str_split(date(), " "), "[", c(5, 2)), collapse = ""),
                   # unique(gfn), # Optional to include names of datasets in output file
                   "_PIO_STRING.txt"),
                 collapse = "")

write.table(x = PPIall, file = fn_ppi, sep = "\t", 
            quote = FALSE, row.names = FALSE, col.names = FALSE)

```


## References

1) [*APID* Alonso-Lopez, D., et al. (2019). "APID database: redefining protein-protein interaction experimental evidences and binary interactomes." Database (Oxford) 2019.](https://pubmed.ncbi.nlm.nih.gov/30715274/)

2) [*APID* Alonso-Lopez, D., et al. (2016). "APID interactomes: providing proteome-based interactomes with controlled quality for multiple species and derived networks." Nucleic Acids Res 44(W1): W529-535.](https://pubmed.ncbi.nlm.nih.gov/27131791/)

3) [*BIOGRID* Oughtred, R., et al. (2021). "The BioGRID database: A comprehensive biomedical resource of curated protein, genetic, and chemical interactions." Protein Sci 30(1): 187-200.](https://pubmed.ncbi.nlm.nih.gov/33070389/)

4) [*BIOPLEX* Huttlin, E. L., et al. (2021). "Dual proteome-scale networks reveal cell-specific remodeling of the human interactome." Cell 184(11): 3022-3040 e3028.](https://pubmed.ncbi.nlm.nih.gov/33961781/)

5) [*CORUM* Tsitsiridis, G., et al. (2023). "CORUM: the comprehensive resource of mammalian protein complexes-2022." Nucleic Acids Res 51(D1): D539-D545.](https://pubmed.ncbi.nlm.nih.gov/36382402/)

6) [*CORUM* Giurgiu, M., et al. (2019). "CORUM: the comprehensive resource of mammalian protein complexes-2019." Nucleic Acids Res 47(D1): D559-D563.](https://pubmed.ncbi.nlm.nih.gov/30357367/)

7) [*HITPREDICT* Lopez, Y., et al. (2015). "HitPredict version 4: comprehensive reliability scoring of physical protein-protein interactions from more than 100 species." Database (Oxford) 2015.](https://pubmed.ncbi.nlm.nih.gov/26708988/)

8) [*HURI* Luck, K., et al. (2020). "A reference map of the human binary protein interactome." Nature 580(7803): 402-408.](https://pubmed.ncbi.nlm.nih.gov/32296183/)

9) [*INTACT* Orchard, S., et al. (2014). "The MIntAct project--IntAct as a common curation platform for 11 molecular interaction databases." Nucleic Acids Res 42(Database issue): D358-363.](https://pubmed.ncbi.nlm.nih.gov/24234451/)

10) [*MINT* Licata, L., et al. (2012). "MINT, the molecular interaction database: 2012 update." Nucleic Acids Res 40(Database issue): D857-861.](https://pubmed.ncbi.nlm.nih.gov/22096227/)

11) [*REACTOME* Gillespie, M., et al. (2022). "The reactome pathway knowledgebase 2022." Nucleic Acids Res 50(D1): D687-D692.](https://pubmed.ncbi.nlm.nih.gov/34788843/)

12) [*STRING* Szklarczyk, D., et al. (2023). "The STRING database in 2023: protein-protein association networks and functional enrichment analyses for any sequenced genome of interest." Nucleic Acids Res 51(D1): D638-D646.](https://pubmed.ncbi.nlm.nih.gov/36370105/)















